Supplementary material

### 2022-01-31 submission/manuscript.rmd
working on updates following co-authors' comments

All risk-based methods assume that a risk prediction model is available to
assign risk predictions to individual patients. For the simulations we developed
a prediction model internally, using logistic regression including main effects
for all baseline covariates and treatment assignment. Risk predictions for
individual patients were based on treatment assignment to the control arm, that
is setting treatment assignment to 0.

@@

Finally, we considered an adaptive approach using Akaike's Information Criterion
(AIC) for model selection. The candidate models were: a constant treatment
effect model, a model with a linear interaction with treatment and RCS models
with 3, 4 and 5 knots. The extra degrees of freedom were 1 (linear interaction),
2, 3 and 4 (RCS models) for these increasingly complex interactions with the
treatment effect.

@@

@@

In comparison with the true approach, dsicrimination for benefit in the scenario
with a constant relative treatment effect was only slightly lower for the linear
interaction model, but substantially lower for the non-linear RCS approaches
(Figure \ref{fig:discrimination}; panel A). With strong linear or quadratic
deviations from a constant relative treatment effect, all methods discriminated
quite similarly (Figure \ref{fig:discrimination}; panels B-C). In the scenario
with non-monotonic deviations, the constant effect model had much lower
discriminative ability compared to all other methods (median AUC of
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_constant)` 
for the constant effects model,
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_linear)`
for the linear interaction model and
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_rcs_3)`
for the best-performing RCS-3; Figure \ref{fig:discrimination}; panel D). The
adaptive approach was unstable in terms of discrimination for benefit,
especially in the presence of treatment-related harms. With increasing number of
RCS knots, we observed decreasing median values and increasing variability of
the c-for-benefit in all scenarios. When we increased the sample size to 17,000
we observed similar trends, however the performance of all methods was more
stable (Supplement, Figure S4). Finally, when we increased the true prediction
AUC to 0.85 the adaptive approach in the case of non-monotonic deviations was,
again, quite unstable, especially with null or moderate treatment-related harms
(Supplement, Figure S5). In these scenarios, the adaptive approach tended to
select more often the inferior constant treatment effect method (Supplement,
Figure S3)

@@

## Case study
```{r, echo=FALSE, warning=FALSE, message=FALSE}
load(here::here("data/raw/gusto.rda"))
gusto <- gusto %>%
  tibble() %>%
  filter(!is.na(tpa))

@@

RCS-4 and RCS-5 proved to be too flexible in all considered scenarios, as
indicated by higher RMSE, increased variability of discrimination for benefit
and worse calibration of benefit predictions. Even with larger sample sizes and
strong quadratic or non-monotonic deviations from the base case scenario of
constant relative treatment effects, these more flexible restricted cubic
splines did not outperform the simpler RCS-3 These approaches may only be
helpful if we expect more extreme patterns of heterogeneous treatment effects
compared to the quadratic deviations considered here. Considering interactions
in RCS-3 models as the most complex approach often may be reasonable.

@@

Risk-based approaches to predictive HTE estimate treatment benefit as a function
of baseline risk. A limitation of our study is that we assumed treatment benefit
to be a function of baseline risk in the majority of the simulation
scenarios. We also considered constant moderate and strong treatment-related
harms, applied on the absolute scale to expand the range of scenarios in line
with previous work (REF!!). Also, we considered a small number of scenarios with
true treatment-covariate interactions, in which our main conclusions remained
the same (Supplement, XX). Future simulation studies could explore the effect of
more extensive deviations from risk-based treatment effects.

Recent years have seen an increased interest in predictive HTE approaches
focusing on individualized benefit predictions. In our simulations we only
focused on risk-based methods, using baseline risk as a reference in a two-stage
approach to individualizing benefit predictions. However, there is a plethora of
different methods, ranging from treatment effect modeling to tree-based
approaches available in more recent literature [@Athey2019; @Lu2018;@Wager2018].
Simulations are also needed to assess relative performance and define the
settings where these break down or outperform each other.

>>>>

All risk-based methods assume that a risk prediction model is available to
assign risk predictions to individual patients. For the simulations we developed
a prediction model internally on the entire population, using a logistic
regression model with main effects for all baseline covariates and treatment
assignment. Risk predictions for individual patients were derived by setting
treatment assignment to 0. Another common approach is to derive the prediction
model solely on the control patients, however this approach has been shown to
lead to biased benefit predictions [@vanKlaveren2019; @Burke2014; @Abadie2018].

@@

Finally, we considered an *adaptive approach* using Akaike's Information
Criterion (AIC) for model selection. More specifically, for the adaptive
approach we ranked the constant relative treatment effect model, the linear
interaction model, and the RCS models with 3, 4, and 5 knots based on their AIC
and selected the one with the lowest value. The extra degrees of freedom were 1
(linear interaction), 2, 3 and 4 (RCS models) for these increasingly complex
interactions with the treatment effect.

@@

## Empirical illustration

```{r, echo=FALSE, warning=FALSE, message=FALSE}
load(here::here("data/raw/gusto.rda"))
gusto <- gusto %>%
  tibble() %>%
  filter(!is.na(tpa))

@@

In comparison with the true approach, dsicrimination for benefit in the scenario
with a constant relative treatment effect was only slightly lower for the linear
interaction model, but substantially lower for the non-linear RCS approaches
(Figure \ref{fig:discrimination}; panel A). With strong linear or quadratic
deviations from a constant relative treatment effect, all methods discriminated
quite similarly (Figure \ref{fig:discrimination}; panels B-C). In the scenario
with non-monotonic deviations, the constant effect model had much lower
discriminative ability compared to all other methods (median AUC of
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_constant)` 
for the constant effects model,
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_linear)`
for the linear interaction model and
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_rcs_3)`
for the best-performing RCS-3; Figure \ref{fig:discrimination}; panel D). The
adaptive approach was unstable in terms of discrimination for benefit,
especially in the presence of treatment-related harms. With increasing number of
RCS knots, we observed decreasing median values and increasing variability of
the c-for-benefit in all scenarios. When we increased the sample size to 17,000
we observed similar trends, however the performance of all methods was more
stable (Supplement, Figure S4). Finally, when we increased the true prediction
AUC to 0.85 the adaptive approach in the case of non-monotonic deviations was,

again, quite unstable, especially with null or moderate treatment-related harms
(Supplement, Figure S5). In these scenarios, the adaptive approach tended to
select more often the inferior constant treatment effect method (Supplement,
Figure S3)

@@

## Empirical illustration

@@

RCS-4 and RCS-5 proved to be too flexible in all considered scenarios, as
indicated by higher RMSE, increased variability of discrimination for benefit
and worse calibration of benefit predictions. Even with larger sample sizes and
strong quadratic or non-monotonic deviations from the base case scenario of
constant relative treatment effects, these more flexible restricted cubic
splines did not outperform the simpler RCS-3. These approaches may only be
helpful if we expect more extreme patterns of heterogeneous treatment effects
compared to the quadratic deviations considered here. Considering interactions
in RCS-3 models as the most complex approach often may be reasonable.

@@

Risk-based approaches to predictive HTE estimate treatment benefit as a function
of baseline risk. A limitation of our study is that we assumed treatment benefit
to be a function of baseline risk in the majority of the simulation
scenarios. We also considered constant moderate and strong treatment-related
harms, applied on the absolute scale to expand the range of scenarios in line
with previous work [@Glasziou1995]. Also, we considered a small number of
scenarios with true treatment-covariate interactions, in which our main
conclusions remained the same (Supplement, XX). Future simulation studies could
explore the effect of more extensive deviations from risk-based treatment
effects.

In our simulations we only focused on risk-based methods, using baseline risk as
a reference in a two-stage approach to individualizing benefit
predictions. However, there is a plethora of different methods, ranging from
treatment effect modeling to tree-based approaches available in more recent
literature [@Athey2019; @Lu2018;@Wager2018;@powers2018some]. Many of these
methods rely on incorporating treatment-covariate interactions in the prediction
of benefit. An important caveat of such approaches is that they may be prone to
overfitting, thus exaggerating the magnitude of the predicted benefits. In a
wide range of simulation settings, a simpler risk modeling approach was
consistently better calibrated for benefit compared to more complex treatment
effect modelling approaches [@vanKlaveren2019]. Similarly, when SYNTAX score II,
a model developed for identifying patients with complex coronary artery disease
that benefit more from percutaneous coronary intervention or from coronary
artery bypass grafting was redeveloped using fewer treatment-covariate
interactions had better external performance compared to its
predecessor[@farooq2013anatomical; @takahashi2020redevelopment].

### 2022-01-31 submission/manuscript.tex
working on updates following co-authors' comments

All risk-based methods assume that a risk prediction model is available
to assign risk predictions to individual patients. For the simulations
we developed a prediction model internally, using logistic regression
including main effects for all baseline covariates and treatment
assignment. Risk predictions for individual patients were based on
treatment assignment to the control arm, that is setting treatment
assignment to 0.

@@

Fourth, we used \emph{restricted cubic splines} (RCS) to relax the
linearity assumption on the effect of the linear predictor {[}8{]}. We
considered splines with 3 (RCS-3), 4 (RCS-4) and 5 (RCS-5) knots to
compare models with different levels of flexibility.

Finally, we considered an adaptive approach using Akaike's Information
Criterion (AIC) for model selection. The candidate models were: a
constant treatment effect model, a model with a linear interaction with
treatment and RCS models with 3, 4 and 5 knots. The extra degrees of
freedom were 1 (linear interaction), 2, 3 and 4 (RCS models) for these
increasingly complex interactions with the treatment effect.

@@

\[\text{RMSE}=\sqrt{\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2}\]
We compared the discriminative ability of the methods under study using
c-for-benefit {[}9{]}. The c-for-benefit represents the probability that
from two randomly chosen matched patient pairs with unequal observed
benefit, the pair with greater observed benefit also has a higher
predicted benefit. To be able to calculate observed benefit, patients in
each treatment arm are ranked based on their predicted benefit and then
matched 1:1 across treatment arms. \emph{Observed} treatment benefit is
defined as the difference of observed outcomes between the untreated and
the treated patient of each matched patient pair. \emph{Predicted}
benefit is defined as the average of predicted benefit within each
matched patient pair.

We evaluated calibration in a similar manner, using the integrated
calibration index (ICI) for benefit {[}10{]}. The observed benefits are
regressed on the predicted benefits using a locally weighted scatterplot
smoother (loess). The ICI-for-benefit is the average absolute difference
between predicted and smooth observed benefit. Values closer to \(0\)
represent better calibration.

@@

In comparison with the true approach, dsicrimination for benefit in the
scenario with a constant relative treatment effect was only slightly
lower for the linear interaction model, but substantially lower for the
non-linear RCS approaches (Figure \ref{fig:discrimination}; panel A).
With strong linear or quadratic deviations from a constant relative
treatment effect, all methods discriminated quite similarly (Figure
\ref{fig:discrimination}; panels B-C). In the scenario with
non-monotonic deviations, the constant effect model had much lower
discriminative ability compared to all other methods (median AUC of
0.4971 for the constant effects model, 0.5285 for the linear interaction
model and 0.5304 for the best-performing RCS-3; Figure
\ref{fig:discrimination}; panel D). The adaptive approach was unstable
in terms of discrimination for benefit, especially in the presence of
treatment-related harms. With increasing number of RCS knots, we
observed decreasing median values and increasing variability of the
c-for-benefit in all scenarios. When we increased the sample size to
17,000 we observed similar trends, however the performance of all
methods was more stable (Supplement, Figure S4). Finally, when we
increased the true prediction AUC to 0.85 the adaptive approach in the
case of non-monotonic deviations was, again, quite unstable, especially
with null or moderate treatment-related harms (Supplement, Figure S5).
In these scenarios, the adaptive approach tended to select more often
the inferior constant treatment effect method (Supplement, Figure S3)

@@

\hypertarget{case-study}{%
\subsection{Case study}\label{case-study}}

In line with previous analyses {[}11,12{]}, we fitted a logistic
regression model with 6 baseline covariates, i.e.~age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and
the location of MI, to predict 30-day mortality risk. A constant effect
of treatment was included in the model. When deriving risk predictions
for individuals we set the treatment indicator to 0. More information on
model development can be found in the supplement.

@@

Increasing the discriminative ability of the risk model--by increasing
the predictor coefficients of the true risk model--reduced RMSE for all
methods. This increase in discriminative ability translates in higher
variability of predicted risks, which, in turn, allows the considered
methods to better capture absolute treatment benefits. As a consequence,
the increase in discriminative ability of the risk model also led to
higher discrimination between those with low or high benefit (as
reflected in values of c-for-benefit). Even though risk model
performance is very important for the ability of risk-based methods to
predict treatment benefit, prediction model development was outside the
scope of this work and has already been studied extensively
{[}5,13,14{]}.

@@

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. We also considered constant moderate and
strong treatment-related harms, applied on the absolute scale to expand
the range of scenarios in line with previous work (REF!!). Also, we
considered a small number of scenarios with true treatment-covariate
interactions, in which our main conclusions remained the same
(Supplement, XX). Future simulation studies could explore the effect of
more extensive deviations from risk-based treatment effects.

Recent years have seen an increased interest in predictive HTE
approaches focusing on individualized benefit predictions. In our
simulations we only focused on risk-based methods, using baseline risk
as a reference in a two-stage approach to individualizing benefit
predictions. However, there is a plethora of different methods, ranging
from treatment effect modeling to tree-based approaches available in
more recent literature {[}15--17{]}. Simulations are also needed to
assess relative performance and define the settings where these break
down or outperform each other.

@@

\leavevmode\hypertarget{ref-Harrell1988}{}%
{[}8{]} Harrell FE, Lee KL, Pollock BG. Regression models in clinical
studies: Determining relationships between predictors and response. JNCI
Journal of the National Cancer Institute 1988;80:1198--202.
\url{https://doi.org/10.1093/jnci/80.15.1198}.

\leavevmode\hypertarget{ref-vanKlaveren2018}{}%
{[}9{]} Klaveren D van, Steyerberg EW, Serruys PW, Kent DM. The proposed
``concordance-statistic for benefit'' provided a useful metric when
modeling heterogeneous treatment effects. Journal of Clinical
Epidemiology 2018;94:59--68.
\url{https://doi.org/10.1016/j.jclinepi.2017.10.021}.

\leavevmode\hypertarget{ref-Austin2019}{}%
{[}10{]} Austin PC, Steyerberg EW. The integrated calibration index
(ICI) and related metrics for quantifying the calibration of logistic
regression models. Statistics in Medicine 2019;38:4051--65.
\url{https://doi.org/10.1002/sim.8281}.

\leavevmode\hypertarget{ref-Califf1997}{}%
{[}11{]} Califf RM, Woodlief LH, Harrell FE, Lee KL, White HD, Guerci A,
et al. Selection of thrombolytic therapy for individual patients:
Development of a clinical model. American Heart Journal 1997;133:630--9.
\url{https://doi.org/10.1016/s0002-8703(97)70164-9}.

\leavevmode\hypertarget{ref-Steyerberg2000}{}%
{[}12{]} Steyerberg EW, Bossuyt PMM, Lee KL. Clinical trials in acute
myocardial infarction: Should we adjust for baseline characteristics?
American Heart Journal 2000;139:745--51.
\url{https://doi.org/10.1016/s0002-8703(00)90001-2}.

\leavevmode\hypertarget{ref-Burke2014}{}%
{[}13{]} Burke JF, Hayward RA, Nelson JP, Kent DM. Using internally
developed risk models to assess heterogeneity in treatment effects in
clinical trials. Circulation: Cardiovascular Quality and Outcomes
2014;7:163--9. \url{https://doi.org/10.1161/circoutcomes.113.000497}.

\leavevmode\hypertarget{ref-Abadie2018}{}%
{[}14{]} Abadie A, Chingos MM, West MR. Endogenous stratification in
randomized experiments. The Review of Economics and Statistics
2018;100:567--80. \url{https://doi.org/10.1162/rest_a_00732}.

@@

\leavevmode\hypertarget{ref-Wager2018}{}%
{[}17{]} Wager S, Athey S. Estimation and inference of heterogeneous
treatment effects using random forests. Journal of the American
Statistical Association 2018;113:1228--42.
\url{https://doi.org/10.1080/01621459.2017.1319839}.
\end{cslreferences}

>>>>

All risk-based methods assume that a risk prediction model is available
to assign risk predictions to individual patients. For the simulations
we developed a prediction model internally on the entire population,
using a logistic regression model with main effects for all baseline
covariates and treatment assignment. Risk predictions for individual
patients were derived by setting treatment assignment to 0. Another
common approach is to derive the prediction model solely on the control
patients, however this approach has been shown to lead to biased benefit
predictions {[}5,8,9{]}.

@@

Fourth, we used \emph{restricted cubic splines} (RCS) to relax the
linearity assumption on the effect of the linear predictor {[}10{]}. We
considered splines with 3 (RCS-3), 4 (RCS-4) and 5 (RCS-5) knots to
compare models with different levels of flexibility.

Finally, we considered an \emph{adaptive approach} using Akaike's
Information Criterion (AIC) for model selection. More specifically, for
the adaptive approach we ranked the constant relative treatment effect
model, the linear interaction model, and the RCS models with 3, 4, and 5
knots based on their AIC and selected the one with the lowest value. The
extra degrees of freedom were 1 (linear interaction), 2, 3 and 4 (RCS
models) for these increasingly complex interactions with the treatment
effect.

@@

\[\text{RMSE}=\sqrt{\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2}\]
We compared the discriminative ability of the methods under study using
c-for-benefit {[}11{]}. The c-for-benefit represents the probability
that from two randomly chosen matched patient pairs with unequal
observed benefit, the pair with greater observed benefit also has a
higher predicted benefit. To be able to calculate observed benefit,
patients in each treatment arm are ranked based on their predicted
benefit and then matched 1:1 across treatment arms. \emph{Observed}
treatment benefit is defined as the difference of observed outcomes
between the untreated and the treated patient of each matched patient
pair. \emph{Predicted} benefit is defined as the average of predicted
benefit within each matched patient pair.

We evaluated calibration in a similar manner, using the integrated
calibration index (ICI) for benefit {[}12{]}. The observed benefits are
regressed on the predicted benefits using a locally weighted scatterplot
smoother (loess). The ICI-for-benefit is the average absolute difference
between predicted and smooth observed benefit. Values closer to \(0\)
represent better calibration.

\hypertarget{empirical-illustration}{%
\subsection{Empirical illustration}\label{empirical-illustration}}

@@

In comparison with the true approach, dsicrimination for benefit in the
scenario with a constant relative treatment effect was only slightly
lower for the linear interaction model, but substantially lower for the
non-linear RCS approaches (Figure \ref{fig:discrimination}; panel A).
With strong linear or quadratic deviations from a constant relative
treatment effect, all methods discriminated quite similarly (Figure
\ref{fig:discrimination}; panels B-C). In the scenario with
non-monotonic deviations, the constant effect model had much lower
discriminative ability compared to all other methods (median AUC of
0.4971 for the constant effects model, 0.5285 for the linear interaction
model and 0.5304 for the best-performing RCS-3; Figure
\ref{fig:discrimination}; panel D). The adaptive approach was unstable
in terms of discrimination for benefit, especially in the presence of
treatment-related harms. With increasing number of RCS knots, we
observed decreasing median values and increasing variability of the
c-for-benefit in all scenarios. When we increased the sample size to
17,000 we observed similar trends, however the performance of all
methods was more stable (Supplement, Figure S4). Finally, when we
increased the true prediction AUC to 0.85 the adaptive approach in the
case of non-monotonic deviations was,

again, quite unstable, especially with null or moderate
treatment-related harms (Supplement, Figure S5). In these scenarios, the
adaptive approach tended to select more often the inferior constant
treatment effect method (Supplement, Figure S3)

@@

\hypertarget{empirical-illustration-1}{%
\subsection{Empirical illustration}\label{empirical-illustration-1}}

In line with previous analyses {[}13,14{]}, we fitted a logistic
regression model with 6 baseline covariates, i.e.~age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and
the location of MI, to predict 30-day mortality risk. A constant effect
of treatment was included in the model. When deriving risk predictions
for individuals we set the treatment indicator to 0. More information on
model development can be found in the supplement.

@@

Increasing the discriminative ability of the risk model--by increasing
the predictor coefficients of the true risk model--reduced RMSE for all
methods. This increase in discriminative ability translates in higher
variability of predicted risks, which, in turn, allows the considered
methods to better capture absolute treatment benefits. As a consequence,
the increase in discriminative ability of the risk model also led to
higher discrimination between those with low or high benefit (as
reflected in values of c-for-benefit). Even though risk model
performance is very important for the ability of risk-based methods to
predict treatment benefit, prediction model development was outside the
scope of this work and has already been studied extensively {[}5,8,9{]}.

@@

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. We also considered constant moderate and
strong treatment-related harms, applied on the absolute scale to expand
the range of scenarios in line with previous work {[}{\textbf{???}}{]}.
Also, we considered a small number of scenarios with true
treatment-covariate interactions, in which our main conclusions remained
the same (Supplement, XX). Future simulation studies could explore the
effect of more extensive deviations from risk-based treatment effects.

In our simulations we only focused on risk-based methods, using baseline
risk as a reference in a two-stage approach to individualizing benefit
predictions. However, there is a plethora of different methods, ranging
from treatment effect modeling to tree-based approaches available in
more recent literature {[}15--18{]}. Many of these methods rely on
incorporating treatment-covariate interactions in the prediction of
benefit. An important caveat of such approaches is that they may be
prone to overfitting, thus exaggerating the magnitude of the predicted
benefits. In a wide range of simulation settings, a simpler risk
modeling approach was consistently better calibrated for benefit
compared to more complex treatment effect modelling approaches {[}5{]}.
Similarly, when SYNTAX score II, a model developed for identifying
patients with complex coronary artery disease that benefit more from
percutaneous coronary intervention or from coronary artery bypass
grafting was redeveloped using fewer treatment-covariate interactions
had better external performance compared to its predecessor{[}19,20{]}.

@@

\leavevmode\hypertarget{ref-Burke2014}{}%
{[}8{]} Burke JF, Hayward RA, Nelson JP, Kent DM. Using internally
developed risk models to assess heterogeneity in treatment effects in
clinical trials. Circulation: Cardiovascular Quality and Outcomes
2014;7:163--9. \url{https://doi.org/10.1161/circoutcomes.113.000497}.

\leavevmode\hypertarget{ref-Abadie2018}{}%
{[}9{]} Abadie A, Chingos MM, West MR. Endogenous stratification in
randomized experiments. The Review of Economics and Statistics
2018;100:567--80. \url{https://doi.org/10.1162/rest_a_00732}.

\leavevmode\hypertarget{ref-Harrell1988}{}%
{[}10{]} Harrell FE, Lee KL, Pollock BG. Regression models in clinical
studies: Determining relationships between predictors and response. JNCI
Journal of the National Cancer Institute 1988;80:1198--202.
\url{https://doi.org/10.1093/jnci/80.15.1198}.

\leavevmode\hypertarget{ref-vanKlaveren2018}{}%
{[}11{]} Klaveren D van, Steyerberg EW, Serruys PW, Kent DM. The
proposed ``concordance-statistic for benefit'' provided a useful metric
when modeling heterogeneous treatment effects. Journal of Clinical
Epidemiology 2018;94:59--68.
\url{https://doi.org/10.1016/j.jclinepi.2017.10.021}.

\leavevmode\hypertarget{ref-Austin2019}{}%
{[}12{]} Austin PC, Steyerberg EW. The integrated calibration index
(ICI) and related metrics for quantifying the calibration of logistic
regression models. Statistics in Medicine 2019;38:4051--65.
\url{https://doi.org/10.1002/sim.8281}.

\leavevmode\hypertarget{ref-Califf1997}{}%
{[}13{]} Califf RM, Woodlief LH, Harrell FE, Lee KL, White HD, Guerci A,
et al. Selection of thrombolytic therapy for individual patients:
Development of a clinical model. American Heart Journal 1997;133:630--9.
\url{https://doi.org/10.1016/s0002-8703(97)70164-9}.

\leavevmode\hypertarget{ref-Steyerberg2000}{}%
{[}14{]} Steyerberg EW, Bossuyt PMM, Lee KL. Clinical trials in acute
myocardial infarction: Should we adjust for baseline characteristics?
American Heart Journal 2000;139:745--51.
\url{https://doi.org/10.1016/s0002-8703(00)90001-2}.

@@

\leavevmode\hypertarget{ref-Wager2018}{}%
{[}17{]} Wager S, Athey S. Estimation and inference of heterogeneous
treatment effects using random forests. Journal of the American
Statistical Association 2018;113:1228--42.
\url{https://doi.org/10.1080/01621459.2017.1319839}.

\leavevmode\hypertarget{ref-powers2018some}{}%
{[}18{]} Powers S, Qian J, Jung K, Schuler A, Shah NH, Hastie T, et al.
Some methods for heterogeneous treatment effect estimation in high
dimensions. Statistics in Medicine 2018;37:1767--87.

\leavevmode\hypertarget{ref-farooq2013anatomical}{}%
{[}19{]} Farooq V, Van Klaveren D, Steyerberg EW, Meliga E, Vergouwe Y,
Chieffo A, et al. Anatomical and clinical characteristics to guide
decision making between coronary artery bypass surgery and percutaneous
coronary intervention for individual patients: Development and
validation of syntax score ii. The Lancet 2013;381:639--50.

\leavevmode\hypertarget{ref-takahashi2020redevelopment}{}%
{[}20{]} Takahashi K, Serruys PW, Fuster V, Farkouh ME, Spertus JA,
Cohen DJ, et al. Redevelopment and validation of the syntax score ii to
individualise decision making between percutaneous and surgical
revascularisation in patients with complex coronary artery disease:
Secondary analysis of the multicentre randomised controlled syntaxes
trial with external cohort validation. The Lancet 2020;396:1399--412.
\end{cslreferences}

### 2022-01-31 submission/supplement.rmd
working on updates following co-authors' comments

```{r tab1, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
 asd <- read_csv(
  here("data/processed/analysisIds.csv")
) %>%
  select(-"type")
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable   = TRUE,
  align       = "r",
  booktabs    = TRUE,
  caption     = "Scenario settings of the entire simulation study.",
  col.names   = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    paste0("g", 0:2),
    "c",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 4,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 7)
```

@@

Here we present the root mean squared error of the considered methods
when true interactions of the covariates with treatment are
assumed. We ran our analyses on a limited set of scenarios. More
specifically, true prediction AUC was set to 0.75 and sample size was
set to 4,250. The simulation scenarios included 4 weak interactions
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.82$), 4 strong interactions
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.61$), and 2 weak and 2
strong interactions.

```{r rmseInteraction, cache=FALSE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replcations calculated in a simulated sample of size 500,000. AUC is 0.85 rather than in Figure \\ref{fig:rmseHighBase}.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_interaction.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

>>>>

```{r tab1, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
asd <- read_csv(
  here("data/processed/analysisIds.csv")
) %>%
  select(-"type")
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable   = TRUE,
  align       = "r",
  booktabs    = TRUE,
  caption     = "Scenario settings of the entire simulation study.",
  col.names   = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    paste0("g", 0:2),
    "c",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 3,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 4,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 7)
```

@@

We carried out a smaller set of simulations, in which we assumed true
treatment-covariate interactions. Sample size was set to 4,250 and the AUC of
the true prediction model was set to 0.75. The following scenarios were
considered: 1) 4 true weak positive interactions
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.83$); 2) 4 strong positive
interactions ($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.61$); 3) 2 weak and 2
strong positive interactions; 4) 4 weak negative interactions 
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=1.17$); 5) 4 strong negative interactions 
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=1.39$); 6) 2 weak and 2 strong negative
interactions; 7) combined positive and negative strong interactions. We also
considered constant treatment-related harms applied on the absolute scale to all
treated patients. The exact settings were: 1) absent treatment-related harms; 2)
moderate treatment-related harms, defined as 25\% of the average true benefit of the
scenario without treatment-related harms; 3) strong treatment-related harms
defined as 50\% of the true average benefit of the scenario without
treatment-related harms; 4) negative treatment-related harms (benefit), defined
as an absolute risk reduction for treated patients of 50\% of the true average
benefit of the scenario without treatment-related harms. The exact settings can
be found in Table \@ref(tab:tab2).

```{r tab2, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
asd <- read_csv(
  here("data/processed/analysisIdsInteractions.csv")
) %>%
  select(-"type")
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable   = TRUE,
  align       = "r",
  booktabs    = TRUE,
  caption     = "Scenario settings of the entire simulation study.",
  col.names   = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    "g0", "g1", "g2", "g5", "g6",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 5,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 6)
```

```{r rmseInteractionPositive, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replcations calculated in a simulated sample of size 500,000. AUC is 0.85 rather than in Figure \\ref{fig:rmseHighBase}.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_interaction_positive.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```
```{r rmseInteractionNegative, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replcations calculated in a simulated sample of size 500,000. AUC is 0.85 rather than in Figure \\ref{fig:rmseHighBase}.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_interaction_negative.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```
```{r rmseInteractionCombined, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replcations calculated in a simulated sample of size 500,000. AUC is 0.85 rather than in Figure \\ref{fig:rmseHighBase}.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_interaction_combined.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

# Empirical illustration

```{r, echo=FALSE, warning=FALSE, message=FALSE}
load(here::here("data/raw/gusto.rda"))
gusto <- gusto %>%
  tibble() %>%
  filter(!is.na(tpa))

gusto <- gusto %>%
  tibble() %>%
  filter(tx != "SK+tPA") %>%
  rename(
    "outcome" = "day30",
    "treatment" = "tpa"
  )

treatmentArms <- gusto %>%
  group_by(treatment) %>%
  summarise(n = n())
```

\begin{equation}
P(outcome=1|X=x) = \text{expit}(lp(x)),
(\#eq:gusto1)
\end{equation}
where 
\begin{equation}
\begin{aligned}
lp(x)=& \beta_0 + \beta_1 \text{age} + \beta_2 I(\text{Killip}=II) + \beta_3I(\text{Killip}=III) +\\
&\beta_4 I(\text{Killip}=IV) + \beta_5ceil(\text{sysbp}, 120) + lsp(\text{pulse}, 50) +\\
&\beta_6 I(\text{pmi}=yes)+\beta_8 I(\text{miloc}=Anterior) + \\
&\beta_9 I(\text{miloc}=Other) + \gamma t_x
\end{aligned}
(\#eq:gusto2)
\end{equation}
